---
title: "TSA_Final_Project"
author: "Kishor Kumar Reddy Mannur"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) 
```

```{r include = FALSE}
setwd("C:/Users/saikr/Desktop/Q4/Time Series/Final Project")
```

```{r}
#Load necessary libraries for time series analyses
library(fpp)        # forecasting principles and practice
library(TSA)        # time series analysis
library(tseries)    # time series analysis and computational finance
library(ggplot2)    # plotting
library(forecast)   # for analyzing uni-variate time series forecasts
library(vars)       # estimation, lag selection, diagnostic testing, causality analyses
library(stats)      # statistical analyses
library(dplyr)      # for data manipulations (filter, arrange, mutate & summarize)
library(xts)
library(lubridate)
```

```{r}
#Reading the climate data
weather_data <- read.csv("weather_data.csv")
head(weather_data)
summary(weather_data)
```

```{r}
#EDA
#Converting to desired datetime format
#weather_data$datetime <- as.POSIXct(weather_data$datetime, format = "%m/%d/%Y")
#weather_data$datetime <- as.POSIXct(weather_data$datetime)
head(weather_data$datetime)
```
```{r}
#write.csv(weather_data, "File_12_Modified.csv", row.names = FALSE)
```


```{r}
library(ggplot2)

# Convert datetime column to POSIXct if needed
weather_data$datetime <- as.POSIXct(weather_data$datetime)

# Filter data for the desired year range (1970 to 2023)
filtered_data <- subset(weather_data, year(datetime) >= 1970 & year(datetime) <= 2023)
#
## Create a ggplot plot
#ggplot(data = filtered_data, aes(x = datetime, y = temp)) +
#  geom_line() +
#  scale_x_datetime(date_labels = "%Y", date_breaks = "1 year") +
#  labs(x = 'Year', y = 'Temperature', title = "Weather Dataset (1970 to 2023)")
```

````{r}
weather_filtered_ts <- ts(filtered_data$temp, start = c(1980, 1), end = c(1983, 12), frequency = 365)
head(weather_filtered_ts)
```

```{r}
plot(weather_filtered_ts, xlab = 'Year', ylab = 'Temperature', main = "Temp Recorded")
```

```{r}
#Checking for seasonality and trend in the dataset
tsdisplay(weather_filtered_ts, main='Temperature 1970 to 2023')
```

### We can see that the ACF plot follows a sine wave patterns which is indicative of seasonality in the dataset. Also the PACF at lag 0 is 1 and cuts off at 0.5 at lag 1. Using a few techniques, we can remove the seasonal component from the data.

```{r}
#Checking for ideal lambda value to transform the series using boxcox
#for (i in -5:5){
#  boxcox <- BoxCox(weather_filtered_ts, lambda = i)
#  print(autoplot(boxcox, main = paste("Box-Cox with Lambda =", i)))
#}
```

```{r}
lambda <- BoxCox.lambda(weather_filtered_ts)
weather_bc <- BoxCox(weather_filtered_ts, BoxCox.lambda(weather_filtered_ts))
plot(weather_bc, main = paste("Transformed Weather Data with Auto Lambda = ", lambda))
```

```{r}
adf_test <- adf.test(weather_filtered_ts)
print(adf_test)
```
### ADF test shows that our data is non-stationary

```{r}
# Load the necessary library
library(trend)

# Perform Mann-Kendall test
mann_kendall_test <- mk.test(weather_filtered_ts)
print(mann_kendall_test)
```
### We can see that the p-value suggests there is no trend in the data, therefore, we proceed with the data as is and apply our models and forecast the data.

```{r}
# Fit SARIMA model
sarima_model <- auto.arima(weather_filtered_ts, seasonal = TRUE)

# Forecast using the SARIMA model
forecast_sarima <- forecast(sarima_model, h = 12)  # Forecast for the next 12 time points

# Plot the SARIMA forecast
plot(forecast_sarima, main = "SARIMA Forecast")
```


```{r}
# Fit ETS model
ets_model <- stlf(weather_filtered_ts)

# Forecast using the ETS model
forecast_ets <- forecast(ets_model, h = 12)  # Forecast for the next 12 time points

# Plot the ETS forecast
plot(forecast_ets, main = "ETS Forecast")
```


```{r}
head(weather_data)
```
```{r}
#Removing NA values
weather_data <- weather_data[complete.cases(weather_data), ]

# Assuming your data is named 'weather_data'
# Convert datetime column to POSIXct format
weather_data$datetime <- as.POSIXct(weather_data$datetime)
head(weather_data)
```

````{r}
# Load the necessary packages
library(prophet)
library(dplyr)

# Create a data frame suitable for Prophet
prophet_data <- weather_data %>%
  select(ds = datetime, y = temp)

# Fit a Prophet model
prophet_model <- prophet(prophet_data, yearly.seasonality = TRUE, daily.seasonality=TRUE)

# Make future dataframe for forecasting
future <- make_future_dataframe(prophet_model, periods = 365)  # Forecast for 365 days

# Forecast using the model
forecast <- predict(prophet_model, future)

# Plot the forecast
plot(prophet_model, forecast)
```