---
title: "Final Project ARIMA monthly"
output: html_notebook
author: "Sai Krishna"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) 
```

```{r include = FALSE}
setwd("C:/Users/saikr/Desktop/Q4/Time Series/Final Project")
```

```{r}
library(forecast)
library(tseries)
library(ggplot2)
library(fpp)
library(tidyverse)
```


```{r}
weather_data <- read.csv("final_data_1970_2023.csv")
weather_data$avgTemp = (weather_data$tempmax + weather_data$tempmin)/2
df_ts <- ts(weather_data, start = c(1970, 1), frequency = 365.25)
plot(df_ts[,'temp'])
checkresiduals(df_ts[,'temp'])
```

```{r}
df_mon <- weather_data %>%
  mutate(weather_data, yearmon = str_sub(weather_data[,'datetime'], start = 1, end = 7) ) %>%
  group_by(yearmon)　%>%
  mutate(avgMTemp = mean(temp)) %>% 
  distinct(yearmon,.keep_all = TRUE) %>%
  select(c(yearmon, avgMTemp))
df_mon_ts = ts(df_mon, start = c(1970, 1), end = c(2023,7), frequency = 12)

df_mon_train <- filter(df_mon, yearmon < "2021-06")
df_mon_test <- filter(df_mon, yearmon >= "2021-06")

# Split the train and test datasets

df_mon_train.ts <-ts(df_mon_train, start = c(1970, 1), end = c(2021,5), frequency = 12)
plot(df_mon_train.ts[,'avgMTemp'])
checkresiduals(df_mon_train.ts[,'avgMTemp'])

df_mon_test.ts <-ts(df_mon_test, start = c(2021,6), end = c(2023,8), frequency = 12)
plot(df_mon_test.ts[,'avgMTemp'], main = "Monthly Average Temparature")
checkresiduals(df_mon_test.ts[,'avgMTemp'])
```
## Check for stationarity
```{r}
Acf(df_mon_ts[,'avgMTemp'])
Pacf(df_mon_ts[,'avgMTemp'])
```
ACF and PACF suggests the time series is auto regressive

```{r}
kpss.test(df_mon_ts[,'avgMTemp'])
adf.test(df_mon_ts[,'avgMTemp'])
```
## The tests do say the data is stationary but visual inspection says it is not

```{r}
BoxCox.lambda(df_mon_ts[,'avgMTemp'])
```
#Use lambda = 1.5 from the following steps

```{r}
df_mon_ts_boxcox <- BoxCox(df_ts[,'temp'], lambda = 1.5)
checkresiduals(df_mon_ts_boxcox)
df_mon_ts_boxcox_d1 <- diff(df_mon_ts_boxcox, differences = 1)
checkresiduals(df_mon_ts_boxcox_d1)
```

### ARIMA

```{r}
mon.fit_arima <- auto.arima(df_mon_ts[,'avgMTemp'], seasonal = TRUE, D = 1, lambda = 1.5, trace = TRUE)
```


```{r}
summary(mon.fit_arima)
checkresiduals(mon.fit_arima)
```

```{r}
source('eacf.R')
```

```{r}
eacf(df_mon_ts[,'avgMTemp'])
```
#Try out ARIMA(3,1,3) as well as per ther EACF recommendation
```{r}
mon.fit.arima102 <- Arima(df_mon_train.ts[,'avgMTemp'], order = c(1,0,2), 
                      seasonal = list(order = c(1,1,0), period = 12), lambda =1.5)
```


```{r}
summary(mon.fit.arima102)
```
## Arima with (1,0,2)(1,1,0) and seasonality 12 is better than the auto-arima suggested model.


```{r}
fcst.mon.arima <- forecast(mon.fit.arima102, h = 25, level = c(80,95))
# plot(fcst.mon.arima)
# Create a plot with forecasted and actual values
plot <- autoplot(fcst.mon.arima) + autolayer(df_mon_test.ts[,'avgMTemp'])
# Specify the x-axis limits for the last 4 years
last_year <- 2018
plot <- plot + scale_x_continuous(
  limits = c(last_year, 2024),
  breaks = seq(last_year, 2024, by = 1)
)
```

```{r}
plot
```

```{r}
accuracy(fcst.mon.arima, df_mon_test.ts[,'avgMTemp'])
```

## Comparing with HW, ARIMA did not outperform the Holt Winters Smoothing in terms of RMSE and MAE.






