---
title: "Final Project ARIMA"
output: html_notebook
author: "Sai Krishna"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) 
```

```{r include = FALSE}
setwd("C:/Users/saikr/Desktop/Q4/Time Series/Final Project")
```

```{r}
library(forecast)
library(tseries)
library(ggplot2)
library(fpp)
library(tidyverse)
```


```{r}
weather_data <- read.csv("final_data_1970_2023.csv")
weather_data$avgTemp = (weather_data$tempmax + weather_data$tempmin)/2
df_ts <- ts(weather_data, start = c(1970, 1), frequency = 365.25)
plot(df_ts[,'temp'])
checkresiduals(df_ts[,'temp'])
```
```{r}
start_date <- as.Date("2000-01-01")
end_date <- as.Date("2023-07-31")

subset_data <- subset(weather_data, datetime >= start_date & datetime <= end_date)
df_ts <- ts(subset_data, start = c(2000, 1), frequency = 365.25)
plot(df_ts[,'temp'])
checkresiduals(df_ts[,'temp'])
```

```{r}
df_ts[,'avgTemp'] <- df_ts[,'avgTemp']+50
df_ts[,'tempmax'] <- df_ts[,'tempmax']+50
df_ts[,'tempmin'] <- df_ts[,'tempmin']+50
df_ts[,'temp'] <- df_ts[,'temp']+50
```

```{r}
plot(df_ts[,'avgTemp'], main = "Daily Average Temparature")
plot(df_ts[,'tempmax'], main = "Daily Max Temparature")
plot(df_ts[,'tempmin'], main = "Daily Min Temparature")
plot(df_ts[,'temp'], main = "Daily Temparature")
```
## Check for stationarity
```{r}
Acf(df_ts[,'temp'])
Pacf(df_ts[,'temp'])
```
ACF and PACF suggests the time series is auto regressive

```{r}
kpss.test(df_ts[,'temp'])
adf.test(df_ts[,'temp'])
```
## The tests do say the data is stationary but visual inspection says it is not

```{r}
BoxCox.lambda(df_ts[,'temp'])
```
#Use lambda = 2 from the following steps

```{r}
df_ts_boxcox <- BoxCox(df_ts[,'temp'], lambda = 2)
checkresiduals(df_ts_boxcox)
df_ts_boxcox_d1 <- diff(df_ts_boxcox, differences = 1)
checkresiduals(df_ts_boxcox_d1)
```

### ARIMA

```{r}
fit_arima <- auto.arima(df_ts[,'temp'], seasonal = TRUE, D = 1, lambda = 2, trace = TRUE)
```


```{r}
summary(fit_arima)
checkresiduals(fit_arima)
```

```{r}
source('eacf.R')
```

```{r}
eacf(df_ts[,'temp'])
```
#Try out ARIMA(3,1,3) as well as per ther EACF recommendation
```{r}
fit.arima313 <- Arima(df_ts[,'temp'], order = c(3,0,3), 
                      seasonal = list(order = c(0,1,0), period = 365.25), lambda =2)
```


```{r}
summary(fit.arima313)
```
## Arima with (3,0,3)(2,1,0) and seasonality 12 is better than the auto-arima suggested model.
However, the AICc is a little high for the model

```{r}
fit.arima.300 <- Arima(df_ts[,'temp'], order = c(3,0,0), 
                      seasonal = list(order = c(0,1,0), period = 365.25), lambda =2)
```

```{r}
summary(fit.arima.300)
```
```{r}
checkresiduals(fit.arima313)
checkresiduals(fit.arima.300)
```

### SARIMA

```{r}
df_temp_ts <- ts(df_ts, start = c(2000, 1), frequency = 365.25)
# Box Cox
df_reg_ts_bc <- df_temp_ts
df_reg_ts_bc[,"temp"] <- BoxCox(df_temp_ts[,"temp"], lambda = 2)

y_day21 <- as.numeric(as.POSIXlt("2021-05-31")$yday)
train_temp <- window(df_reg_ts_bc, start = c(2000, 1), end = c(2021, y_day21))
train_temp2 <- window(df_temp_ts, start = c(2000, 1), end = c(2021, y_day21))

yday_23 <- as.numeric(as.POSIXlt("2023-07-31")$yday)
temp_test <- window(df_reg_ts_bc, start = c(2021, y_day21 + 1), end = c(2023, yday_23))
temp_test2 <- window(df_temp_ts, start = c(2021, y_day21 + 1), end = c(2023, yday_23))
```


```{r}
fit.sarima<- auto.arima(train_temp2[,'temp'], seasonal = TRUE, trace = TRUE, 
                        max.p = 3,max.q = 3, max.P = 2, max.Q = 2,
                         allowdrift = FALSE, num.cores = 10)
```

```{r}
summary(fit.sarima)
checkresiduals(fit.sarima)
```

```{r}
df2 <- data.frame(time = time(pred$pred), value = as.numeric(pred$pred))
test_df2 <- data.frame(time = time(temp_test2[,'temp']), value = as.numeric(temp_test2[,'temp']))
```

```{r}
df.ts.2 <- ts(df2, start = c(2021,410), frequency = 365)
df.ts.2.test <- ts(test_df2, start = c(2021, 410), frequency = 365)

```


```{r}
# pred = predict(fit.sarima, n.ahead = 790)
accuracy(df.ts.2, df.ts.2.test)
```

```{r}
# Create a plot with forecasted and actual values
plot <- autoplot(forecast(fit.sarima, h = 790, level = c(80,95))) #+ autolayer(df.ts.2)

# Specify the x-axis limits for the last 4 years
last_year <- 2018
plot <- plot + scale_x_continuous(
  limits = c(last_year, 2024),
  breaks = seq(last_year, 2024, by = 1)
)

# Display the plot
print(plot)
```


```{r}
plot(forecast(fit.sarima, h = 365, level = c(80,95)))
```
```{r}
summary(fit.sarima)
checkresiduals(fit.sarima)
```
## SARIMA did better in AICc



